---
title: "Code for analyses"
output:
  word_document: default
  html_document: default
date: "2024-04-15"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning = FALSE)
```

This is the code used to produce the results in Cahill & Chenuil.

```{r}
library(dplyr)
library(stringr)
library(ggplot2)
library(vioplot)
library(bestglm)
library(tidyr)
library(ggmosaic)
library(lme4)
library(cowplot)
library(GLMMselect)
library(cAIC4)
library(wesanderson)

survey <- read.csv(file="978_species_clean_may17.csv" , header=TRUE ) 
bio<-read.csv2(file="input_BIOLOGY_classes.csv")

survey <-as.data.frame(unclass(survey),stringsAsFactors=TRUE)

```


# Analyses with full dataset (starting n = 978)

1) Are there more CSss or pseudo-CS overall, considering only cases where morphology was measured? (n = 492)

```{r}
expected<-c((length(na.omit(survey$Morpho_diff)))/2,(length(na.omit(survey$Morpho_diff)))/2)
observed<-c(length(which(survey$CSss=="TRUE")),length(which(survey$CSss=="FALSE")))
chisq.test(cbind(expected,observed))
```

2) Are there differences across phyla or classes? (n = 492)
```{r}
surveymorpho<-filter(survey, Morpho_diff != "NA") #make matrix where only the cases where morpho was studied is here

drop1(glm(CSss~phylum_wormsV1,data=surveymorpho,family=binomial),test="Chisq")
drop1(glm(CSss~phylum_class,data=surveymorpho,family=binomial),test="Chisq")

```

Since phylum_class is significant, we will now use this as a random factor in subsequent models.

3) Is the probability of CSss associated with year of description? (n = 492)

```{r}
summary(glmer(CSss~yearb+(1|phylum_class),data=surveymorpho,family=binomial))

```

4) Is there a relationship between the number of CS in easily-accessible or inaccessible habitats and time? (n = 972)

```{r}
surveyaccess<-filter(survey, HDifAccess != "NA")
surveyaccess$Difficult<-surveyaccess$HDifAccess=="Inaccessible"
surveyaccess$Easy<-surveyaccess$HDifAccess=="Accessible"

summary(glmer(Difficult~yearb+(1|phylum_class),data=surveyaccess,family=binomial))
summary(glmer(Easy~yearb+(1|phylum_class),data=surveyaccess,family=binomial))

```

5) Is habitat connectivity (variable HKK) associated with CSss? (n = 492)

```{r}
summary(glmer(CSss~HKKv3+(1|phylum_class),data=surveymorpho,family=binomial))

```

6) Are CSss more common in rocky or soft sediment habitats? (n = 227) What about when only considering shallow-water species? (n = 193)
```{r}
surveysed<-filter(surveymorpho, Hsubstrate != "NA") #make matrix where only the cases for rocky or sediment habitat
surveysed<-filter(surveysed, Hsubstrate != "Sediment and Rocky") #and remove the cases with both

summary(glmer(CSss~Hsubstrate+(1|phylum_class),data=surveysed,family=binomial))

shallow<-filter(surveymorpho, shallow == "TRUE")
shallow<-filter(shallow, Hsubstrate != "Sediment and Rocky") #and remove the cases with both, 192 cases remaining

summary(glmer(CSss~Hsubstrate+(1|phylum_class),data=shallow,family=binomial))


```

7) Are CSss more common in sympatry or no? (n = 486)
```{r}
summary(glmer(CSss~Sympatric+(1|phylum_class),data=surveymorpho,family=binomial))
```

8) Are species in sympatry more likely to have ecological differentiation? (n=978)
```{r}
summary(glmer(Eco_diff~Sympatric+(1|phylum_class),data=survey,family=binomial))

```

9) Are CSss more common in species with ecological differentiation? (n = 107)
```{r}
summary(glmer(CSss~Eco_diff+(1|phylum_class),data=surveymorpho,family=binomial))

```

10) Are biological traits (hard skeleton, fertilization type, presence of external genitalia, image-forming vision) associated with CSss? (n = 464)
```{r}
summary(glmer(CSss~hard_skeleton+(1|phylum_class),data=surveymorpho,family=binomial))
summary(glmer(CSss~ferti+(1|phylum_class),data=surveymorpho,family=binomial))
summary(glmer(CSss~genitals+(1|phylum_class),data=surveymorpho,family=binomial))
summary(glmer(CSss~image+(1|phylum_class),data=surveymorpho,family=binomial))

```

11) Is larval type associated with CSss? (n = 305)
```{r}
surveylarvmorpho<-filter(surveymorpho, Larv_type != "NA")

summary(glmer(CSss~Larv_type+(1|phylum_class),data=surveylarvmorpho,family=binomial))
```

12) Are CSss more common in particular geographic zones? (n = 367)
```{r}
tropmorpho<-filter(surveymorpho, Trop != "NA") 
summary(glmer(CSss~Trop+(1|phylum_class),data=tropmorpho,family=binomial))

npolmorpho<-filter(surveymorpho, Npol != "NA") 
summary(glmer(CSss~Npol+(1|phylum_class),data=npolmorpho,family=binomial))

ntempmorpho<-filter(surveymorpho, Ntemp != "NA") 
summary(glmer(CSss~Ntemp+(1|phylum_class),data=ntempmorpho,family=binomial))

spolmorpho<-filter(surveymorpho, Spol != "NA") #make matrix where only the cases where morpho was studied is here
summary(glmer(CSss~Spol+(1|phylum_class),data=spolmorpho,family=binomial))

stempmorpho<-filter(surveymorpho, Stemp != "NA") 
summary(glmer(CSss~Stemp+(1|phylum_class),data=stempmorpho,family=binomial))

```

13) What about the bestglm to determine which factors are most important?

First with all data
```{r}
mat<-select(surveymorpho, c(HKKv3, phylum_class, Sympatric,CSss))
mat[1:4]<-lapply(mat[1:4],as.factor)
outBIC<-bestglm(mat,IC="BIC", family=binomial,TopModels=10)  
outBIC$BestModels 

```

And now with the shallow-water rocky vs soft sed dataset
```{r}
shallow<-filter(shallow, Hsubstrate != "Sediment and Rocky") #and remove the cases  with both

mat<-select(shallow, c(Hsubstrate, HKKv3, phylum_class, Sympatric, CSss))
mat[1:5]<-lapply(mat[1:5],as.factor)
outBIC<-bestglm(mat,IC="BIC", family=binomial,TopModels=10)  
outBIC$BestModels 

```


# Analyses with species where reproductive isolation is clear based on multiple datatypes (see main text) (starting n = 535)

```{r}
survey <- read.csv(file="full_dataset" , header=TRUE ) 
bio<-read.csv2(file="input_BIOLOGY_classes.csv")
surveylarv<-read.csv("survey_oct11.csv",header=TRUE)

survey <-as.data.frame(unclass(survey),stringsAsFactors=TRUE)
survey<-filter(survey, RIOK==TRUE) #N = 535
surveylarv<-filter(surveylarv, RIOK==TRUE) 

```
1) Are there more CSss or pseudo-CS overall, considering only cases where morphology was measured? (n = 411)

```{r}
expected<-c((length(na.omit(survey$Morpho_diff)))/2,(length(na.omit(survey$Morpho_diff)))/2)
observed<-c(length(which(survey$CSss=="TRUE")),length(which(survey$CSss=="FALSE")))
chisq.test(cbind(expected,observed))
```

2) Are there differences across phyla or classes? (n = 411)
```{r}
surveymorpho<-filter(survey, Morpho_diff != "NA") #make matrix where only the cases where morpho was studied is here

drop1(glm(CSss~phylum_wormsV1,data=surveymorpho,family=binomial),test="Chisq")
drop1(glm(CSss~phylum_class,data=surveymorpho,family=binomial),test="Chisq")

```

Since phylum_class is significant, we will now use this as a random factor in subsequent models.

3) Is the probability of CSss associated with year of description? (n = 347)

```{r}
summary(glmer(CSss~yearb+(1|phylum_class),data=surveymorpho,family=binomial))

```

4) Is there a relationship between the number of CS in easily-accessible or inaccessible habitats and time? (n = 450)

```{r}
surveyaccess<-filter(survey, HDifAccess != "NA")
surveyaccess$Difficult<-surveyaccess$HDifAccess=="Inaccessible"
surveyaccess$Easy<-surveyaccess$HDifAccess=="Accessible"

summary(glmer(Difficult~yearb+(1|phylum_class),data=surveyaccess,family=binomial))
summary(glmer(Easy~yearb+(1|phylum_class),data=surveyaccess,family=binomial))

```

5) Is habitat connectivity (variable HKK) associated with CSss? (n = 378)

```{r}
summary(glmer(CSss~HKKv3+(1|phylum_class),data=surveymorpho,family=binomial))

```

6) Are CSss more common in rocky or soft sediment habitats? (n = 190) What about when only considering shallow-water species? (n = 190)
```{r}
surveysed<-filter(surveymorpho, Hsubstrate != "NA") #make matrix where only the cases for rocky or sediment habitat
surveysed<-filter(surveysed, Hsubstrate != "Sediment and Rocky") #and remove the cases with both

summary(glmer(CSss~Hsubstrate+(1|phylum_class),data=surveysed,family=binomial))

shallow<-read.csv("shallow_rock_sed.csv",header=T)
shallow<-filter(shallow, Hsubstrate != "Sediment and Rocky") #and remove the cases with both

summary(glmer(CSss~Hsubstrate+(1|phylum_class),data=shallow,family=binomial))


```

7) Are CSss more common in sympatry or no? (n = 407)
```{r}
summary(glmer(CSss~Sympatric+(1|phylum_class),data=surveymorpho,family=binomial))
```

8) Are species in sympatry more likely to have ecological differentiation? (n=140)
```{r}
summary(glmer(Eco_diff~Sympatric+(1|phylum_class),data=survey,family=binomial))

```

9) Are CSss more common in species with ecological differentiation? (n = 104)
```{r}
summary(glmer(CSss~Eco_diff+(1|phylum_class),data=surveymorpho,family=binomial))

```

10) Are biological traits (hard skeleton, fertilization type, presence of external genitalia, image-forming vision) associated with CSss? (n = 386)
```{r}
summary(glmer(CSss~hard_skeleton+(1|phylum_class),data=surveymorpho,family=binomial))
summary(glmer(CSss~ferti+(1|phylum_class),data=surveymorpho,family=binomial))
summary(glmer(CSss~genitals+(1|phylum_class),data=surveymorpho,family=binomial))
summary(glmer(CSss~image+(1|phylum_class),data=surveymorpho,family=binomial))

```

11) Is larval type associated with CSss? (n = 246)
```{r}
surveylarvmorpho<-filter(surveylarv, Morpho_diff != "NA")
surveylarvmorpho<-filter(surveylarvmorpho, Larv_type != "NA")

summary(glmer(CSss~Larv_type+(1|phylum_class),data=surveylarvmorpho,family=binomial))
```

12) Are CSss more common in particular geographic zones? (n = 312)
```{r}
tropmorpho<-filter(surveymorpho, Trop != "NA") 
summary(glmer(CSss~Trop+(1|phylum_class),data=tropmorpho,family=binomial))

npolmorpho<-filter(surveymorpho, Npol != "NA") 
summary(glmer(CSss~Npol+(1|phylum_class),data=npolmorpho,family=binomial))

ntempmorpho<-filter(surveymorpho, Ntemp != "NA") 
summary(glmer(CSss~Ntemp+(1|phylum_class),data=ntempmorpho,family=binomial))

spolmorpho<-filter(surveymorpho, Spol != "NA") 
summary(glmer(CSss~Spol+(1|phylum_class),data=spolmorpho,family=binomial))

stempmorpho<-filter(surveymorpho, Stemp != "NA") 
summary(glmer(CSss~Stemp+(1|phylum_class),data=stempmorpho,family=binomial))

```


# Analyses with species where the number of CS per article = 1 (see main text) (starting n = 535)

```{r}
survey <- read.csv(file="full_dataset" , header=TRUE ) 
bio<-read.csv2(file="input_BIOLOGY_classes.csv")
surveylarv<-read.csv("survey_oct11.csv",header=TRUE)

survey<-filter(survey, Nb_CScomplex_inArticle == 1) 
surveylarv<-filter(surveylarv, Nb_CScomplex_inArticle == 1)

```
1) Are there more CSss or pseudo-CS overall, considering only cases where morphology was measured? (n = 395)

```{r}
expected<-c((length(na.omit(survey$Morpho_diff)))/2,(length(na.omit(survey$Morpho_diff)))/2)
observed<-c(length(which(survey$CSss=="TRUE")),length(which(survey$CSss=="FALSE")))
chisq.test(cbind(expected,observed))
```

2) Are there differences across phyla or classes? (n = 395)
```{r}
surveymorpho<-filter(survey, Morpho_diff != "NA") #make matrix where only the cases where morpho was studied is here

drop1(glm(CSss~phylum_wormsV1,data=surveymorpho,family=binomial),test="Chisq")
drop1(glm(CSss~phylum_class,data=surveymorpho,family=binomial),test="Chisq")

```

Since phylum_class is significant, we will now use this as a random factor in subsequent models.

3) Is the probability of CSss associated with year of description? (n = 324)

```{r}
summary(glmer(CSss~yearb+(1|phylum_class),data=surveymorpho,family=binomial))

```

4) Is there a relationship between the number of CS in easily-accessible or inaccessible habitats and time? (n = 395)

```{r}
surveyaccess<-filter(survey, HDifAccess != "NA")
surveyaccess$Difficult<-surveyaccess$HDifAccess=="Inaccessible"
surveyaccess$Easy<-surveyaccess$HDifAccess=="Accessible"

summary(glmer(Difficult~yearb+(1|phylum_class),data=surveyaccess,family=binomial))
summary(glmer(Easy~yearb+(1|phylum_class),data=surveyaccess,family=binomial))

```

5) Is habitat connectivity (variable HKK) associated with CSss? (n = 360)

```{r}
summary(glmer(CSss~HKKv3+(1|phylum_class),data=surveymorpho,family=binomial))

```

6) Are CSss more common in rocky or soft sediment habitats? (n = 171) What about when only considering shallow-water species? (n = 171)
```{r}
surveysed<-filter(surveymorpho, Hsubstrate != "NA") #make matrix where only the cases for rocky or sediment habitat
surveysed<-filter(surveysed, Hsubstrate != "Sediment and Rocky") #and remove the cases with both

summary(glmer(CSss~Hsubstrate+(1|phylum_class),data=surveysed,family=binomial))

shallow<-read.csv("shallow_rock_sed.csv",header=T)
shallow<-filter(shallow, Hsubstrate != "Sediment and Rocky") #and remove the cases with both

summary(glmer(CSss~Hsubstrate+(1|phylum_class),data=shallow,family=binomial))


```

7) Are CSss more common in sympatry or no? (n = 390)
```{r}
summary(glmer(CSss~Sympatric+(1|phylum_class),data=surveymorpho,family=binomial))
```

8) Are species in sympatry more likely to have ecological differentiation? (n=125)
```{r}
summary(glmer(Eco_diff~Sympatric+(1|phylum_class),data=survey,family=binomial))

```

9) Are CSss more common in species with ecological differentiation? (n = 91)
```{r}
summary(glmer(CSss~Eco_diff+(1|phylum_class),data=surveymorpho,family=binomial))

```

10) Are biological traits (hard skeleton, fertilization type, presence of external genitalia, image-forming vision) associated with CSss? (n = 368)
```{r}
summary(glmer(CSss~hard_skeleton+(1|phylum_class),data=surveymorpho,family=binomial))
summary(glmer(CSss~ferti+(1|phylum_class),data=surveymorpho,family=binomial))
summary(glmer(CSss~genitals+(1|phylum_class),data=surveymorpho,family=binomial))
summary(glmer(CSss~image+(1|phylum_class),data=surveymorpho,family=binomial))

```

11) Is larval type associated with CSss? (n = 261)
```{r}
surveylarvmorpho<-filter(surveylarv, Morpho_diff != "NA")
surveylarvmorpho<-filter(surveylarvmorpho, Larv_type != "NA")

summary(glmer(CSss~Larv_type+(1|phylum_class),data=surveylarvmorpho,family=binomial))
```

12) Are CSss more common in particular geographic zones? (n = 294)
```{r}
tropmorpho<-filter(surveymorpho, Trop != "NA")
summary(glmer(CSss~Trop+(1|phylum_class),data=tropmorpho,family=binomial))

npolmorpho<-filter(surveymorpho, Npol != "NA")
summary(glmer(CSss~Npol+(1|phylum_class),data=npolmorpho,family=binomial))

ntempmorpho<-filter(surveymorpho, Ntemp != "NA") 
summary(glmer(CSss~Ntemp+(1|phylum_class),data=ntempmorpho,family=binomial))

spolmorpho<-filter(surveymorpho, Spol != "NA") 
summary(glmer(CSss~Spol+(1|phylum_class),data=spolmorpho,family=binomial))

stempmorpho<-filter(surveymorpho, Stemp != "NA") 
summary(glmer(CSss~Stemp+(1|phylum_class),data=stempmorpho,family=binomial))

```

# Analyses with species where the number of individuals studied for genetics is greater than 20 (see main text) (starting n = 735)

```{r}
survey <- read.csv(file="full_dataset" , header=TRUE ) 
bio<-read.csv2(file="input_BIOLOGY_classes.csv")
surveylarv<-read.csv("survey_oct11.csv",header=TRUE)

survey<-filter(survey, as.numeric(survey$Nig) > 20) 

surveylarv <-as.data.frame(unclass(surveylarv),stringsAsFactors=TRUE)
surveylarv<-filter(surveylarv, as.numeric(surveylarv$Nig) > 20)


```
1) Are there more CSss or pseudo-CS overall, considering only cases where morphology was measured? (n = 383)

```{r}
expected<-c((length(na.omit(survey$Morpho_diff)))/2,(length(na.omit(survey$Morpho_diff)))/2)
observed<-c(length(which(survey$CSss=="TRUE")),length(which(survey$CSss=="FALSE")))
chisq.test(cbind(expected,observed))
```

2) Are there differences across phyla or classes? (n = 383)
```{r}
surveymorpho<-filter(survey, Morpho_diff != "NA") #make matrix where only the cases where morpho was studied is here

drop1(glm(CSss~phylum_wormsV1,data=surveymorpho,family=binomial),test="Chisq")
drop1(glm(CSss~phylum_class,data=surveymorpho,family=binomial),test="Chisq")

```

Since phylum_class is significant, we will now use this as a random factor in subsequent models.

3) Is the probability of CSss associated with year of description? (n = 319)

```{r}
summary(glmer(CSss~yearb+(1|phylum_class),data=surveymorpho,family=binomial))

```

4) Is there a relationship between the number of CS in easily-accessible or inaccessible habitats and time? (n = 616)

```{r}
surveyaccess<-filter(survey, HDifAccess != "NA")
surveyaccess$Difficult<-surveyaccess$HDifAccess=="Inaccessible"
surveyaccess$Easy<-surveyaccess$HDifAccess=="Accessible"

summary(glmer(Difficult~yearb+(1|phylum_class),data=surveyaccess,family=binomial))
summary(glmer(Easy~yearb+(1|phylum_class),data=surveyaccess,family=binomial))

```

5) Is habitat connectivity (variable HKK) associated with CSss? (n = 351)

```{r}
summary(glmer(CSss~HKKv3+(1|phylum_class),data=surveymorpho,family=binomial))

```

6) Are CSss more common in rocky or soft sediment habitats? (n = 160) What about when only considering shallow-water species? (n = 160)
```{r}
surveysed<-filter(surveymorpho, Hsubstrate != "NA") #make matrix where only the cases for rocky or sediment habitat
surveysed<-filter(surveysed, Hsubstrate != "Sediment and Rocky") #and remove the cases with both

summary(glmer(CSss~Hsubstrate+(1|phylum_class),data=surveysed,family=binomial))

shallow<-read.csv("shallow_rock_sed.csv",header=T)
shallow<-filter(shallow, Hsubstrate != "Sediment and Rocky") #and remove the cases with both

summary(glmer(CSss~Hsubstrate+(1|phylum_class),data=shallow,family=binomial))


```

7) Are CSss more common in sympatry or no? (n = 379)
```{r}
summary(glmer(CSss~Sympatric+(1|phylum_class),data=surveymorpho,family=binomial))
```

8) Are species in sympatry more likely to have ecological differentiation? (n=114)
```{r}
summary(glmer(Eco_diff~Sympatric+(1|phylum_class),data=survey,family=binomial))

```

9) Are CSss more common in species with ecological differentiation? (n = 81)
```{r}
summary(glmer(CSss~Eco_diff+(1|phylum_class),data=surveymorpho,family=binomial))

```

10) Are biological traits (hard skeleton, fertilization type, presence of external genitalia, image-forming vision) associated with CSss? (n = 361)
```{r}
summary(glmer(CSss~hard_skeleton+(1|phylum_class),data=surveymorpho,family=binomial))
summary(glmer(CSss~ferti+(1|phylum_class),data=surveymorpho,family=binomial))
summary(glmer(CSss~genitals+(1|phylum_class),data=surveymorpho,family=binomial))
summary(glmer(CSss~image+(1|phylum_class),data=surveymorpho,family=binomial))

```

11) Is larval type associated with CSss? (n = 236)
```{r}
surveylarvmorpho<-filter(surveylarv, Morpho_diff != "NA")
surveylarvmorpho<-filter(surveylarvmorpho, Larv_type != "NA")

summary(glmer(CSss~Larv_type+(1|phylum_class),data=surveylarvmorpho,family=binomial))
```

12) Are CSss more common in particular geographic zones? (n = 289)
```{r}
tropmorpho<-filter(surveymorpho, Trop != "NA") 
summary(glmer(CSss~Trop+(1|phylum_class),data=tropmorpho,family=binomial))

npolmorpho<-filter(surveymorpho, Npol != "NA") 
summary(glmer(CSss~Npol+(1|phylum_class),data=npolmorpho,family=binomial))

ntempmorpho<-filter(surveymorpho, Ntemp != "NA") #make matrix where only the cases where morpho was studied is here
summary(glmer(CSss~Ntemp+(1|phylum_class),data=ntempmorpho,family=binomial))

spolmorpho<-filter(surveymorpho, Spol != "NA") 
summary(glmer(CSss~Spol+(1|phylum_class),data=spolmorpho,family=binomial))

stempmorpho<-filter(surveymorpho, Stemp != "NA") 
summary(glmer(CSss~Stemp+(1|phylum_class),data=stempmorpho,family=binomial))

```

