---
title: "Missing CS for paper 3"
output: html_document
date: "2024-06-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## This file is to deal with analyses for paper 3.
They will be in two parts. The first part will be phylogeny, and I want to create a heatmap like in Paper 1. There will be 57 classes on the tree, and 4 columns: number CS in database, number spp in WoRMS, number missing species, number missing bioloigcal species.

I will have the same four variables for a four-panel map figure, based on FAO Marine Regions.

```{r}
library(dplyr)
library(stringr)
library(ggplot2)
library(tidyr)
library(cowplot)
library(wesanderson)
library(ggtree)


survey <- read.csv(file="978_species_clean_may17.csv"   , header=TRUE ) #Adjusted Nov 15 2023 to fix errors in duplicated species
surveyclass<-filter(survey, class_wormsV1 != "NA")
surveyclass2<-filter(surveyclass, acceptedName_wormsV1 !="NA") #824 cases

tree39data<-read.tree("C:/Users/aecsk/Documents/GitHub/Cryptic_species_figs/39classes.tre")
classnames<-read.table("C:/Users/aecsk/Documents/GitHub/Cryptic_species_figs/classnames2024.txt",header=T) # adjust this file so it is only the non-zero classes
nomsp <-read.csv2(file="file_S2_oct.csv" ,header=T)# need to load all nominal spp data 

```

```{r}
# This loop replaces the order names within Hexanauplia so that our database matches the CS output.
for (i in 1:length(nomsp$order)){
  if (nomsp$order[i] == "Calanoida" )
    nomsp$class[i] <- "Copepoda"
  else if (nomsp$order[i] == "Canuelloida" )
    nomsp$class[i] <- "Copepoda"
  else if (nomsp$order[i] == "Cyclopoida" )
    nomsp$class[i] <- "Copepoda"
  else if (nomsp$order[i] == "Harpacticoida" )
    nomsp$class[i] <- "Copepoda"
  else if (nomsp$order[i] == "Misophrioida" )
    nomsp$class[i] <- "Copepoda"
  else if (nomsp$order[i] == "Monstrilloida" )
    nomsp$class[i] <- "Copepoda"
  else if (nomsp$order[i] == "Platycopioida" )
    nomsp$class[i] <- "Copepoda"
  else if (nomsp$order[i] == "Siphonostomatoida" )
    nomsp$class[i] <- "Copepoda"
  else if (nomsp$order[i] == "Cryptophialida" )
    nomsp$class[i] <- "Thecostraca"
  else if (nomsp$order[i] == "Dendrogastrida" )
    nomsp$class[i] <- "Thecostraca"
  else if (nomsp$order[i] == "Ibliformes" )
    nomsp$class[i] <- "Thecostraca"
  else if (nomsp$order[i] == "Laurida" )
    nomsp$class[i] <- "Thecostraca"
  else if (nomsp$order[i] == "Lepadiformes" )
    nomsp$class[i] <- "Thecostraca"
  else if (nomsp$order[i] == "Lithoglyptida" )
    nomsp$class[i] <- "Thecostraca"
  else if (nomsp$order[i] == "Scalpelliformes" )
    nomsp$class[i] <- "Thecostraca"
  else if (nomsp$order[i] == "Sessilia" )
    nomsp$class[i] <- "Thecostraca"
  
  }
```



I may have to remake the tree with just the classes represented in our database???

This makes sense to me because the classes NOT in our database can't tell us anything about missing CS -- they are just ... missing. No comparison to make. Paper 1 used residuals from a larger model, so expected values weren't zeroes. 

37 classes left where we have CS *assigned to species* in our database (all nemerteans lumped for that count)


```{r}

# THERE ARE ZEROES HERE THAT SHOULD NOT BE HERE!  Problems with Thecostraca, Copepoda, Nemertea

prop_missing <- c()
num_CS <- c()
num_worms <- c()
Nb_CS <- c()
surveyclass3<-filter(surveyclass2,Nb_CS != "NA") #799 cases remaining here, look into who I've tossed


for (i in classnames$Class){
  
  d<-nrow(filter(nomsp, (class==i)&(ncbi==T))) #total in class with NCBI
  e<-nrow(filter(nomsp, (class==i))) #total in class
  f<-nrow(filter(nomsp, (class==i)&(hasCS==T))) #total CS
  g<-(f/d)*e #total CS in phylum IF everyone had NCBI
  h<-g-f  #total MISSING CS, yet to be found
  j<-h/g  #proportion of MISSING CS
  k <- filter(surveyclass3,class_wormsV1 == i)
  m <- mean(k$Nb_CS)
  p<-m*h

    num_CS<-rbind(num_CS,f)
  num_worms<-rbind(num_worms,e)
  prop_missing<-rbind(prop_missing,j)
  Nb_CS<-rbind(Nb_CS,p)
}

data_heatmap<-cbind(classnames$Class,num_CS,num_worms,prop_missing,Nb_CS)

#tapply(surveyclass3$Nb_CS,surveyclass3$class_wormsV1,mean)


#for (i in classnames$Class){
 # k <- filter(surveyclass3,class_wormsV1 == i)
 # m <- mean(k$Nb_CS)
 # p<-m*h
 # Nb_CS<-rbind(Nb_CS,p)
#}


write.csv(data_heatmap,"data_for_heatmap2.csv")
```

Now let's make a heatmap

```{r}
traits39<-read.csv("C:/Users/aecsk/Documents/GitHub/Cryptic_species_figs/heatmap_CS3_input.csv",header=T)

supp.labs <- c("Number CS", "Number WoRMS","Proportion Missing","Biological missing")

names(supp.labs) <- c("Number CS", "Number WoRMS","Proportion Missing","Biological missing")

heatmap<-ggplot(data = traits39, mapping = aes(x = Variable,
                                               y = Class_num,
                                               fill = z_score)) +
  geom_tile() +
  scale_fill_gradient2(low = 'blue', mid = 'white', high = 'red')+
  theme_bw()+
  theme(axis.title.x = element_blank(), 
        axis.title.y = element_blank(), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  
        panel.grid.minor = element_blank(),  
        plot.background = element_blank())+
  theme(legend.position="none") +
  theme(axis.text.x=element_blank(),axis.text.y=element_blank())+
  geom_hline(yintercept=3.5)+
  geom_hline(yintercept=7.5)+
  geom_hline(yintercept=12.5)+
  geom_hline(yintercept=13.5)+
  geom_hline(yintercept=18.5)+
  geom_hline(yintercept=20.5)+
  geom_hline(yintercept=21.5)+
  geom_hline(yintercept=28.5)+
  geom_hline(yintercept=29.5)+
  geom_hline(yintercept=33.5)+
  geom_hline(yintercept=34.5)+
  geom_hline(yintercept=36.5)+
  facet_grid(~Variable,labeller = labeller(Variable=supp.labs),switch = "x", scales = "free_x", space = "free_x")


```

and a tree

```{r}
tree39<-ggtree(tree39data)+geom_tiplab()

ggarrange(
    #alldataplot,# First row with line plot
    # Second row with box and dot plots
    tree39, heatmap, 
    ncol = 2) 

```

Note from Jan 2 to make a barplot of missing classes - gode from MissingCS.R file
```{r}
largeclasses<-read.csv("largeclasses.csv",header=TRUE)
ggplot(largeclasses,aes(x=V1,y=as.numeric(V4)))+
  geom_bar(stat="identity")+
  ylab("Proportion CS missing")+
  xlab("Class")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 90))+
  theme(panel.background = element_blank(), 
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.background = element_blank())
```

