---
title: "Annelids only - 133 CS"
output: html_document
date: "2023-10-17"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning = FALSE)
```


## Script to work on analyses for causes paper

```{r}
library(dplyr)
library(stringr)
library(ggplot2)
library(vioplot)
library(bestglm)
library(tidyr)
library(ggmosaic)
survey <- read.csv(file="977_CScpx_100cols_Habitat&Larva_Completed_20230928_comma.csv"   , header=TRUE ) # file ready for analyses with new variables ,# data with range sizes rarefied for 100 sites, and RIOK
nomsp <-read.csv2(file="file_S2_oct.csv" ,header=T)# need to load all nominal spp data for some comparisons 
bio<-read.csv2(file="input_BIOLOGY_classes.csv")
surveylarv<-read.csv("survey_oct11.csv",header=TRUE)


# cleaned from some original variables and temporary variable (CAUTION: CHECK COLUMN nb!):
# reintroduce genus but several possibilities: solution below not OK for CS at genus level (no data), can also use NameBoforeCor instead of acceptedName
survey$genus <-word(survey$acceptedName_wormsV1,1)
survey$genusBC <-word(survey$NameBeforeCor,1) # see if this genusBC changes results / genus.
survey <-as.data.frame(unclass(survey),stringsAsFactors=TRUE)
survey$Geo_dif3 <-survey$Geo_diff
levels(survey$Geo_dif3)<-c("","Allopatric", "Allopatric", "Allopatric","Sympatric","Sympatric, Allopatric", "Sympatric, Allopatric", "Sympatric, Allopatric")
levels(survey$Geo_dif3)<-c(NA, "Allopatric","Sympatric","Sympatric, Allopatric")
# add booleans for geographical differentiation and morphological differentiation variables
survey$symp    <- str_detect(survey$Geo_diff,"Sympatric")
survey$sympara <- str_detect(survey$Geo_diff,"Sympatric") | str_detect(survey$Geo_diff,"Parapatric")
survey$allop   <- str_detect(survey$Geo_diff,"Allopatric")
survey$NoMorphoDiff    <- survey$Morpho_diff=="No differentiation"
survey$StatMorphoDiff  <- survey$Morpho_diff=="Statistic"
survey$DiagMorphoDiff  <- survey$Morpho_diff=="Diagnostic"
# add biological traits of the class to which the CScpx belongs
survey$phylum_class <-paste0(survey$phylum_worms,"_",survey$class_worms)
survey$hard_skeleton  <- factor(bio$hard_skeleton[match(survey$phylum_class,bio$phylum_class)])
survey$ferti          <- factor(bio$ferti[match(survey$phylum_class,bio$phylum_class)])
survey$image          <- factor(bio$image[match(survey$phylum_class,bio$phylum_class)])
survey$genitals       <- factor(bio$genitals[match(survey$phylum_class,bio$phylum_class)])
survey$CSss     <-str_detect(survey$Morpho_diff,"Statistic") | str_detect(survey$Morpho_diff,"No differentiation") #need to combine stat + no_diff to make CSss
survey$Sympatric <-str_detect(survey$Geo_dif3,"Sympatric") | str_detect(survey$Geo_dif3,"Sympatric, Allopatric") #This creates a vector that counts sym/allo as sympatric - that is, do spp exist in sympatry?

#NEW FOR THIS SCRIPT: MAKE RIOK

survey<-filter(survey, phylum_wormsV1=="Annelida")
surveylarv<-filter(surveylarv, phylum_wormsV1=="Annelida")

```
## Taxonomic work is needed? # Morpho differentiation types: More CS sl than CS ss overall? # Need to find expectations and test against 50:50 mark
```{r}
expected<-c((length(na.omit(survey$Morpho_diff)))/2,(length(na.omit(survey$Morpho_diff)))/2)
observed<-c(length(which(survey$CSss=="TRUE")),length(which(survey$CSss=="FALSE")))
chisq.test(cbind(expected,observed))

# Next we will make a graph
CSssGraph<-as.data.frame(cbind(observed,c("CSss","CSsl")))
colnames(CSssGraph)<-c("Number","CS_Type")
ggplot(data=CSssGraph,aes(x=CS_Type, y=Number))+geom_bar(stat="identity")+
  theme_bw()+
  #theme(legend.position="none")+
  theme(panel.background = element_blank(), 
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.background = element_blank())+
  theme(axis.text.x = element_text(angle = 65,hjust=1))
```

More CSss than the 50% benchmark (significant), so reject the idea that most CS are just due to taxonomic oversight.

## Available technology – More diagnostic differences in species described earlier?

```{r}
breaks=c(1750,1760,1770,1780,1790,1800,1810,1820,1830,1840,1850,1860,1870,1880,1890,1900,1910,
         1920,1930,1940,1950,1960,1970,1980,1990,2000,2010,2020,2030)
survey$decade<-cut(as.numeric(survey$yearb),breaks=breaks)
levels(survey$decade)<-c("1750","1760","1770","1780","1790","1800","1810","1820","1830","1840",
                       "1850","1860","1870","1880","1890","1900","1910","1920","1930","1940",
                       "1950","1960","1970","1980","1990","2000","2010","2020","2030")

decademorpho<-cbind(as.character(survey$decade),as.character(survey$CSss))
decademorpho<-(na.omit(decademorpho)) #400 cases left
colnames(decademorpho)<-c("decade","CSss")
decademorpho<-as.data.frame(decademorpho)

table(decademorpho$decade,decademorpho$CSss)
chisq.test(decademorpho$decade,decademorpho$CSss)

decadetable<-table(decademorpho$decade,decademorpho$CSss)
percent_sl<-c()

for (i in 1:length(decadetable[,2])) {
  b<-(decadetable[i,1]/(decadetable[i,1]+decadetable[i,2]))
  percent_sl<-rbind(percent_sl,b)
}

decadetable<-as.data.frame(cbind(decadetable,percent_sl))
colnames(decadetable)<-c("SL","SS","Percent_SL")

decadetable$decade<-c("1760","1770","1800","1810","1820","1830","1840",
                      "1850","1860","1870","1880","1890","1900","1910","1920","1930","1950","1980","1990","2000","2010")

ggplot(decadetable,aes(x=decade,y=Percent_SL))+geom_point()+theme_bw()
```

There is no trend in the percent of CSsl (diagnostic differences) through time; hypothesis not supported.

## Prevailing theories of nature and species origins - More CS in fragmented habitats in species described before Darwin?
```{r}
break3<-c(1756,1859,1980,2020) # periods limited / Darwin, Kimura, then end of survey
survey$y3<-cut(as.numeric(survey$yearb),breaks=break3)
levels(survey$y3)<-c("1756-1859","1860-1980","1981-2020")

breaks_HKK<-cbind(as.character(survey$y3),survey$HKKv3)
breaks_HKK<-(na.omit(breaks_HKK)) #754 cases left
colnames(breaks_HKK)<-c("decade_bin","HKK")
breaks_HKK<-as.data.frame(breaks_HKK)

table(breaks_HKK$decade_bin,breaks_HKK$HKK)
chisq.test(breaks_HKK$decade_bin,breaks_HKK$HKK)

breaks_count<-table(breaks_HKK$decade_bin)
HKKdiff<-table(breaks_HKK$decade_bin,breaks_HKK$HKK) 
pre_Darwin<-HKKdiff[1,]/breaks_count[1]
pre_Kimura<-HKKdiff[2,]/breaks_count[2]
modern<-HKKdiff[3,]/breaks_count[3]
decadebreakstable<-rbind(pre_Darwin,pre_Kimura,modern)
decadebreakstable #this should be the percentages for each, sym and not, columns sum to 1

HKK_decades<-pivot_longer(as.data.frame(decadebreakstable), cols=1:5, names_to = "HKK",
                       values_to = "Percent", values_drop_na = FALSE)
HKK_decades$Era<-c("1Pre_Darwin","1Pre_Darwin","1Pre_Darwin","1Pre_Darwin","1Pre_Darwin","2Pre_Kimura","2Pre_Kimura","2Pre_Kimura","2Pre_Kimura","2Pre_Kimura","3Modern","3Modern","3Modern","3Modern","3Modern")
HKK_decades["HKK"][HKK_decades["HKK"] == "1"] <- "A1"
HKK_decades["HKK"][HKK_decades["HKK"] == "5"] <- "B5"
HKK_decades["HKK"][HKK_decades["HKK"] == "10"] <- "C10"
HKK_decades["HKK"][HKK_decades["HKK"] == "25"] <- "D25"
HKK_decades["HKK"][HKK_decades["HKK"] == "100"] <- "E100"
HKK_decades["HKK"][HKK_decades["HKK"] == "1000"] <- "F1000"

ggplot(as.data.frame(HKK_decades),aes(x=Era,y=Percent,fill=HKK))+geom_bar(stat="identity")
```

This is marginally significant. HKK=5 and 100 both trend UP, and 25 trends DOWN. A bit messy.

## Accessibility of habitat when species described - Percentage of CS in inaccessible habitats higher in earlier spp?

```{r}
decadeAccess<-cbind(as.character(survey$decade),as.character(survey$HDifAccess))
decadeAccess<-(na.omit(decadeAccess)) #832 cases left
colnames(decadeAccess)<-c("decade","Access")
decadeAccess<-as.data.frame(decadeAccess)

table(decadeAccess$decade,decadeAccess$Access)
chisq.test(decadeAccess$decade,decadeAccess$Access)

decadetable<-table(decadeAccess$decade,decadeAccess$Access)
percent_easy<-c()

for (i in 1:length(decadetable[,2])) {
  b<-(decadetable[i,2]/(decadetable[i,1]+decadetable[i,2]+decadetable[i,3]))
  percent_easy<-rbind(percent_easy,b)
}

decadetable<-as.data.frame(cbind(decadetable,percent_easy))
colnames(decadetable)<-c("Difficult","Easy","Normal","Percent_Easy")

decadetable$decade<-c("1750","1760","1770","1780","1790","1800","1810","1820","1830","1840",
                      "1850","1860","1870","1880","1890","1910","1920","1930","1940",
                      "1960","1970","1980","1990","2000","2010")

ggplot(decadetable,aes(x=decade,y=Percent_Easy))+geom_point()+theme_bw()

```

No real trend or support..

## In sum, when we only look at species with CS (i.e. our database), the effect of taxonomic work goes away. This is interesting since that was the strongest effect in paper 1. THIS IS STILL TRUE with the annelids.

## Effect of recent divergence - dispersal

## Larval type – Planktotrophic/pelagic larvae leads to less CS/NS)?
Note1: for this I’m working with a separate surveylarv input file where I have corrected larval types to pt, lt, dd, and removed all others, including cases where there are multiple types checked. (Most of those cases are where the two CS have different larvae.) This left 59 CS with larval info.

Note2: the data fail the homoscedasticity test (Bartlett’s) so I’ve used a nonparametric Kruskal-Wallis test to see if CS/NS is different among groups.
```{r}
larv_NbCS<-cbind(surveylarv$Nb_CS,surveylarv$Larv_type)
larv_NbCS<-as.data.frame(na.omit(larv_NbCS))
colnames(larv_NbCS)<-c("Nb_CS","Larv_type")

#data fail Bartlett's test so kruskal-wallis it is

tapply(as.numeric(larv_NbCS$Nb_CS),larv_NbCS$Larv_type,mean)
tapply(as.numeric(larv_NbCS$Nb_CS),larv_NbCS$Larv_type,sd)
kruskal.test(as.numeric(larv_NbCS$Nb_CS),larv_NbCS$Larv_type)

#graph?
data_summary <- function(x) {
  m <- mean(x)
  ymin <- m-sd(x)
  ymax <- m+sd(x)
  return(c(y=m,ymin=ymin,ymax=ymax))
}
ggplot(data=larv_NbCS,aes(x=Larv_type, y=as.numeric(Nb_CS)))+
  geom_jitter()+
  labs(x ="Larval type", y = "Nb CS per NS")+
  stat_summary(fun.data=data_summary, color="blue",size=0.5)+
  theme_bw()
```

Hypothesis not supported.

##High HKK leads to more CS ss – testing this based on Nb_CS per HKK.
Note: data again failed homoscedasticity, using Kruskal Wallis test
```{r}
HKK_NbCS<-cbind(survey$Nb_CS,survey$HKKv3)
HKK_NbCS<-as.data.frame(na.omit(HKK_NbCS))
colnames(HKK_NbCS)<-c("Nb_CS","HKKv3")

tapply(HKK_NbCS$Nb_CS,HKK_NbCS$HKKv3,mean)
tapply(HKK_NbCS$Nb_CS,HKK_NbCS$HKKv3,sd)
kruskal.test(HKK_NbCS$Nb_CS,HKK_NbCS$HKKv3)

#No difference in the number of CS per NS across HKK classes; the average for all is around 3. 
#graph?
data_summary <- function(x) {
  m <- mean(x)
  ymin <- m-sd(x)
  ymax <- m+sd(x)
  return(c(y=m,ymin=ymin,ymax=ymax))
}
HKK_NbCS["HKKv3"][HKK_NbCS["HKKv3"] == "1"] <- "A1"
HKK_NbCS["HKKv3"][HKK_NbCS["HKKv3"] == "5"] <- "B2"
HKK_NbCS["HKKv3"][HKK_NbCS["HKKv3"] == "10"] <- "C10"
HKK_NbCS["HKKv3"][HKK_NbCS["HKKv3"] == "25"] <- "D25"
HKK_NbCS["HKKv3"][HKK_NbCS["HKKv3"] == "100"] <- "E100"
HKK_NbCS["HKKv3"][HKK_NbCS["HKKv3"] == "1000"] <- "F1000"

ggplot(data=HKK_NbCS,aes(x=HKKv3, y=Nb_CS))+
  geom_jitter()+
  labs(x ="HKK", y = "Nb CS per NS")+
  stat_summary(fun.data=data_summary, color="blue",size=0.5)+
  theme_bw()
```

Result: Hypothesis not supported.

## In sum, NO support for the role of recent divergence as measured by a proxy of dispersal when considering only the annelids.

## Stabilizing selection

## More stable environments = more CSss or statistical differences
Note: stable = pelagic, deepsea, cave, coralreef; not stable = coastal, intertidal, estuarine, seagrass

```{r}
Hstable<-c()

for (i in 1:length(survey$NameBeforeCor)) {
  if (is.na(survey$Hcoralreef[i]))
    b<-"NA"
  else if (survey$Hpelagic[i] == TRUE)
    b<-"TRUE"
  else if (survey$Hdeepsea[i] == TRUE)
    b<-"TRUE"
  else if (survey$Hcave[i] == TRUE)
    b<-"TRUE"
  else if (survey$Hcoralreef[i]  == TRUE)
    b<-"TRUE"
    else
    b<-"FALSE"
  
  Hstable<-c(Hstable,b)
  
}

survey<-cbind(survey,Hstable)

stable_hab<-cbind(survey$Hstable,survey$Morpho_diff,survey$CSss)
stable_hab<-as.data.frame(na.omit(stable_hab)) #477 cases left
colnames(stable_hab)<-c("Stable","Morpho_diff","CSss")

table(stable_hab$CSss,stable_hab$Stable)
chisq.test(stable_hab$CSss,stable_hab$Stable)

stable_count<-table(stable_hab$Stable)
stablediff<-table(stable_hab$CSss,stable_hab$Stable) 
CSss_total<-stablediff[2,]/stable_count
CSsl_total<-stablediff[1,]/stable_count
stabletable<-rbind(CSss_total,CSsl_total)
stabletable #this should be the percentages for each, sym and not, columns sum to 1

surveystabletable<-pivot_longer(as.data.frame(stabletable), cols=1:2, names_to = "Stable",
                               values_to = "Percent", values_drop_na = FALSE)
surveystabletable$CSss_type<-c("CSss","CSss","CSsl","CSsl")

ggplot(as.data.frame(surveystabletable),aes(x=Stable,y=Percent,fill=CSss_type,))+geom_bar(stat="identity")
```

Result: Hypothesis not supported; no differences.

## Note also that stabilizing selection predicts more CSss than CSsl, which is what we find (see above).
## In sum, we generally have some support for this idea.

## High Ne (neutral hypotheses)
##Reproductive investment - Low reproductive investment leads to more statistical divergence? (Repro. investment proxy for Ne)
Note1: Again this uses the surveylarv dataset with hand cleaning. 
Note2: Planktotrophic = LOW; Lecithotrophic, Brooding, DD = HIGH; If a spp could be both low and high, have marked it NA (is usually bc the CS have different larval types). Anything else got marked NA to be conservative. 
Note3: Only 29 cases are considered here, meaning that is where we have morphological data AND larval type data AND Annelid

```{r}
table(surveylarv$Morpho_diff,surveylarv$Repro_invest)
chisq.test(surveylarv$Morpho_diff,surveylarv$Repro_invest)

```

Result: not significant.

## High HKK leads to more CS ss (HKK as proxy for Ne)

```{r}
surveymorpho<-filter(survey, Morpho_diff != "NA") #make matrix where only the cases where morpho was studied is here
Hkk_count<-table(surveymorpho$HKKv3)
morphodiff<-table(surveymorpho$Morpho_diff,surveymorpho$HKKv3) 
stat_diag<-morphodiff[3,]/morphodiff[1,]
stat_anydiff<-morphodiff[3,]/(morphodiff[3,]+morphodiff[1,])
stat_total<-morphodiff[3,]/Hkk_count
diag_total<-morphodiff[1,]/Hkk_count
nod_total<-morphodiff[2,]/Hkk_count
HKKtable<-rbind(stat_total,diag_total,nod_total)
morphodiff 
HKKtable
Hkk_count
chisq.test(surveymorpho$Morpho_diff,surveymorpho$HKKv3)

#Graph
HKKtable2<-pivot_longer(as.data.frame(HKKtable), cols=1:5, names_to = "HKK",
                        values_to = "Percent", values_drop_na = FALSE)
HKKtable2$Diff_type<-c("Statistical","Statistical","Statistical","Statistical","Statistical","Diagnostic","Diagnostic","Diagnostic","Diagnostic","Diagnostic","No_Difference","No_Difference","No_Difference","No_Difference","No_Difference")
HKKtable2["HKK"][HKKtable2["HKK"] == "1"] <- "A1"
HKKtable2["HKK"][HKKtable2["HKK"] == "5"] <- "B2"
HKKtable2["HKK"][HKKtable2["HKK"] == "10"] <- "C10"
HKKtable2["HKK"][HKKtable2["HKK"] == "25"] <- "D25"
HKKtable2["HKK"][HKKtable2["HKK"] == "100"] <- "E100"
HKKtable2["HKK"][HKKtable2["HKK"] == "1000"] <- "F1000"

ggplot(as.data.frame(HKKtable2),aes(x=HKK,y=Percent,fill=Diff_type,))+geom_bar(stat="identity")

#Ok, what if we just consider CSss vs CSsl?

table(surveymorpho$HKKv3,surveymorpho$CSss)
chisq.test(surveymorpho$HKKv3,surveymorpho$CSss)

CSssdiff<-table(surveymorpho$CSss,surveymorpho$HKKv3) 
CSss_total<-CSssdiff[2,]/Hkk_count
CSsl_total<-CSssdiff[1,]/Hkk_count
HKKtable3<-rbind(CSss_total,CSsl_total)
HKKtable3

#Graph
HKK_NbCS<-pivot_longer(as.data.frame(HKKtable3), cols=1:5, names_to = "HKK",
                        values_to = "Percent", values_drop_na = FALSE)
HKK_NbCS$Diff_type<-c("CSss","CSss","CSss","CSss","CSss","CSsl","CSsl","CSsl","CSsl","CSsl")
HKK_NbCS["HKK"][HKK_NbCS["HKK"] == "1"] <- "A1"
HKK_NbCS["HKK"][HKK_NbCS["HKK"] == "5"] <- "B2"
HKK_NbCS["HKK"][HKK_NbCS["HKK"] == "10"] <- "C10"
HKK_NbCS["HKK"][HKK_NbCS["HKK"] == "25"] <- "D25"
HKK_NbCS["HKK"][HKK_NbCS["HKK"] == "100"] <- "E100"
HKK_NbCS["HKK"][HKK_NbCS["HKK"] == "1000"] <- "F1000"

ggplot(as.data.frame(HKK_NbCS),aes(x=HKK,y=Percent,fill=Diff_type,))+geom_bar(stat="identity")
```

Results: The results are significant when considering 3 classes, with more no-diff at higher HKK values and more statistical at lower values, but when considering just ss vs sl, difference goes away. So, overall, hypothesis not supported.

## In sum, we do not see support for the neutral hypothesis in the annelids.

## Hypothesis: effect of recent divergence – More sympatric than allopatric CS/NS for a given Nsg?

```{r}
geo_NbCS<-as.data.frame(cbind(survey$Nb_CS,survey$Sympatric,survey$Nsg))
geo_NbCS<-(na.omit(geo_NbCS))
colnames(geo_NbCS)<-c("Nb_CS","Sympatric","Nsg")
geo_NbCS["Sympatric"][geo_NbCS["Sympatric"] == "1"] <- "TRUE"
geo_NbCS["Sympatric"][geo_NbCS["Sympatric"] == "0"] <- "FALSE"


ggplot(geo_NbCS,aes(x=Nsg,y=Nb_CS,colour=Sympatric))+geom_point()+geom_smooth(method=lm,se=T, fullrange=TRUE)

#So this looks like it tracks with the assumption, but let's test with ANCOVA
library(rstatix)
res.aov <- geo_NbCS %>% anova_test(Nsg ~ Nb_CS + Sympatric)
res.aov
```

These results aren’t statistically significant (ANCOVA) but the graph is in line with the hypothesis. Marginal support. NOTE though that there’s a significant interaction between smpatry and the slope of the line! Not sure how to interpret.

## Selection against hybrids – More diagnostic differences in sympatry than allopatry?

```{r}
table(surveymorpho$CSss,surveymorpho$Sympatric)
chisq.test(surveymorpho$CSss,surveymorpho$Sympatric)

CSsym_count<-table(surveymorpho$Sympatric)
CSssdiff<-table(surveymorpho$CSss,surveymorpho$Sympatric) 
CSss_total<-CSssdiff[2,]/CSsym_count
CSsl_total<-CSssdiff[1,]/CSsym_count
CSsymtable<-rbind(CSss_total,CSsl_total)
CSsymtable #this should be the percentages for each, sym and not, columns sum to 1

surveyCSsymtable<-pivot_longer(as.data.frame(CSsymtable), cols=1:2, names_to = "Sympatry",
                                   values_to = "Percent", values_drop_na = FALSE)
surveyCSsymtable$CSss_type<-c("CSsl","CSsl","CSss","CSss")

ggplot(as.data.frame(surveyCSsymtable),aes(x=Sympatry,y=Percent,fill=CSss_type,))+geom_bar(stat="identity")
```
Result: hypothesis not supported, but v small sample sizes here.

## Importance of biological traits

# More diagnostic traits with hard skeleton?

```{r}
surveymorpho<-filter(survey, Morpho_diff != "NA") #make matrix where only the cases where morpho was studied is here
table(surveymorpho$Morpho_diff,surveymorpho$hard_skeleton)
#chisq.test(surveymorpho$Morpho_diff,surveymorpho$hard_skeleton)

#graph
morphoskel_count<-table(surveymorpho$hard_skeleton)
morphodiff<-table(surveymorpho$Morpho_diff,surveymorpho$hard_skeleton) 
stat_diag<-morphodiff[3,]/morphodiff[1,]
stat_anydiff<-morphodiff[3,]/(morphodiff[3,]+morphodiff[1,])
stat_total<-morphodiff[3,]/morphoskel_count
diag_total<-morphodiff[1,]/morphoskel_count
nod_total<-morphodiff[2,]/morphoskel_count
morphoskeltable<-rbind(stat_total,diag_total,nod_total)
morphoskeltable #this should be the percentages for each, 
#columns sum to 1

surveymorphoskeltable<-pivot_longer(as.data.frame(morphoskeltable), cols=1:2, names_to = "Hard_Skeleton",
                                    values_to = "Percent", values_drop_na = FALSE)
surveymorphoskeltable$Diff_type<-c("Statistical","Statistical","Diagnostic","Diagnostic","No_Difference","No_Difference")

ggplot(as.data.frame(surveymorphoskeltable),aes(x=Hard_Skeleton,y=Percent,fill=Diff_type,))+geom_bar(stat="identity")
```

Nobody has a hard skeleton, so no test.

## More diagnostic traits with image vision?
```{r}
#Next: image vision
table(surveymorpho$Morpho_diff,surveymorpho$image)
#chisq.test(surveymorpho$Morpho_diff,surveymorpho$image)

morphoimage_count<-table(surveymorpho$image)
morphodiff<-table(surveymorpho$Morpho_diff,surveymorpho$image) 
stat_diag<-morphodiff[3,]/morphodiff[1,]
stat_anydiff<-morphodiff[3,]/(morphodiff[3,]+morphodiff[1,])
stat_total<-morphodiff[3,]/morphoimage_count
diag_total<-morphodiff[1,]/morphoimage_count
nod_total<-morphodiff[2,]/morphoimage_count
morphoimagetable<-rbind(stat_total,diag_total,nod_total)
morphoimagetable #this should be the percentages for each, 
#columns sum to 1

surveymorphoimagetable<-pivot_longer(as.data.frame(morphoimagetable), cols=1:2, names_to = "Image_Vision",
                                    values_to = "Percent", values_drop_na = FALSE)
surveymorphoimagetable$Diff_type<-c("Statistical","Statistical","Diagnostic","Diagnostic","No_Difference","No_Difference")

ggplot(as.data.frame(surveymorphoimagetable),aes(x=Image_Vision,y=Percent,fill=Diff_type,))+geom_bar(stat="identity")
```

Result: No image vision, so no test.

## More diagnostic traits with internal fertilization?
```{r}
table(surveymorpho$Morpho_diff,surveymorpho$ferti)
chisq.test(surveymorpho$Morpho_diff,surveymorpho$ferti)

#Note: Still significant if we just do ss vs sl. Graphing whole thing.
morphoferti_count<-table(surveymorpho$ferti)
morphodiff<-table(surveymorpho$Morpho_diff,surveymorpho$ferti) 
stat_diag<-morphodiff[3,]/morphodiff[1,]
stat_anydiff<-morphodiff[3,]/(morphodiff[3,]+morphodiff[1,])
stat_total<-morphodiff[3,]/morphoferti_count
diag_total<-morphodiff[1,]/morphoferti_count
nod_total<-morphodiff[2,]/morphoferti_count
morphofertitable<-rbind(stat_total,diag_total,nod_total)
morphofertitable #this should be the percentages for each, eco and not
#columns sum to 1
surveymorphofertitable<-pivot_longer(as.data.frame(morphofertitable), cols=1:3, names_to = "Fertilization",
                                    values_to = "Percent", values_drop_na = FALSE)
surveymorphofertitable$Diff_type<-c("Statistical","Statistical","Statistical","Diagnostic","Diagnostic","Diagnostic","No_Difference","No_Difference","No_Difference")

ggplot(as.data.frame(surveymorphofertitable),aes(x=Fertilization,y=Percent,fill=Diff_type,))+geom_bar(stat="identity")

```
Result: hypothesis not supported. VERY small number (3) with external fertilization; most are ‘both’.

## More diagnostic traits with external genitalia?

```{r}
table(surveymorpho$Morpho_diff,surveymorpho$genitals)
#chisq.test(surveymorpho$Morpho_diff,surveymorpho$genitals)

#and a graph

morphogen_count<-table(surveymorpho$genitals)
morphodiff<-table(surveymorpho$Morpho_diff,surveymorpho$genitals) 
stat_diag<-morphodiff[3,]/morphodiff[1,]
stat_anydiff<-morphodiff[3,]/(morphodiff[3,]+morphodiff[1,])
stat_total<-morphodiff[3,]/morphogen_count
diag_total<-morphodiff[1,]/morphogen_count
nod_total<-morphodiff[2,]/morphogen_count
morphogentable<-rbind(stat_total,diag_total,nod_total)
morphogentable #this should be the percentages for each, eco and not
#columns sum to 1

surveymorphogentable<-pivot_longer(as.data.frame(morphogentable), cols=1:3, names_to = "External_genitals",
                                   values_to = "Percent", values_drop_na = FALSE)
surveymorphogentable$Diff_type<-c("Statistical","Statistical","Statistical","Diagnostic","Diagnostic","Diagnostic","No_Difference","No_Difference","No_Difference")

ggplot(as.data.frame(surveymorphogentable),aes(x=External_genitals,y=Percent,fill=Diff_type,))+geom_bar(stat="identity")
```

Result: No test, everyone lacks genitalia.

## In sum, biological traits don’t really have a role here - they’re mostly fixed in the phylum.

## So, overall conclusions: much like the cnidarians, there’s not support for very much here. Might be because of low sample size but might ALSO be because there’s just really high percentages of CS in this group (Cnidarians and Annelids are among the highest percentage) so we can’t pick out a signal beyond that.