---
title: "Testing Morphological Differentiation against various factors"

output: html_document
date: "2023-10-05"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning = FALSE)
```

```{r}
library(dplyr)
library(stringr)
library(ggplot2)
library(vioplot)
library(bestglm)
library(tidyr)
library(ggmosaic)

survey <- read.csv(file="977_CScpx_100cols_Habitat&Larva_Completed_20230928_comma.csv"   , header=TRUE ) # file ready for analyses with new variables ,# data with range sizes rarefied for 100 sites, and RIOK
#survey <- read.csv2(file="670spp_65var_HKKv3_2023-01-17.csv"   , header=TRUE ) # file ready for analyses with new variables ,# data with range sizes rarefied for 100 sites, and RIOK
nomsp <-read.csv2(file="file_S2_oct.csv" ,header=T)# need to load all nominal spp data for some comparisons 
bio<-read.csv2(file="input_BIOLOGY_classes.csv")

# cleaned from some original variables and temporary variable (CAUTION: CHECK COLUMN nb!):
#survey <-select(survey, -c(2:9))  # <- Abby commented this out because the old and new versions of 'survey' do not have the same number of columns in the same order. WORK WITH NAMES.
# reintroduce genus but several possibilities: solution below not OK for CS at genus level (no data), can also use NameBoforeCor instead of acceptedName
survey$genus <-word(survey$acceptedName_wormsV1,1)
survey$genusBC <-word(survey$NameBeforeCor,1) # see if this genusBC changes results / genus.
survey <-as.data.frame(unclass(survey),stringsAsFactors=TRUE)
survey$Geo_dif3 <-survey$Geo_diff
levels(survey$Geo_dif3)<-c("","Allopatric", "Allopatric", "Allopatric","Sympatric","Sympatric, Allopatric", "Sympatric, Allopatric", "Sympatric, Allopatric")
levels(survey$Geo_dif3)<-c(NA, "Allopatric","Sympatric","Sympatric, Allopatric")
# add booleans for geographical differentiation and morphological differentiation variables
survey$symp    <- str_detect(survey$Geo_diff,"Sympatric")
survey$sympara <- str_detect(survey$Geo_diff,"Sympatric") | str_detect(survey$Geo_diff,"Parapatric")
survey$allop   <- str_detect(survey$Geo_diff,"Allopatric")
survey$NoMorphoDiff    <- survey$Morpho_diff=="No differentiation"
survey$StatMorphoDiff  <- survey$Morpho_diff=="Statistic"
survey$DiagMorphoDiff  <- survey$Morpho_diff=="Diagnostic"
# add biological traits of the class to which the CScpx belongs
survey$phylum_class <-paste0(survey$phylum_worms,"_",survey$class_worms)
survey$hard_skeleton  <- factor(bio$hard_skeleton[match(survey$phylum_class,bio$phylum_class)])
survey$ferti          <- factor(bio$ferti[match(survey$phylum_class,bio$phylum_class)])
survey$image          <- factor(bio$image[match(survey$phylum_class,bio$phylum_class)])
survey$genitals       <- factor(bio$genitals[match(survey$phylum_class,bio$phylum_class)])


```

## Big Question 1: Hypotheses of what causes CS (related to Ne)
#1. We have three classes of CS: no difference (CSss), Stat difference, and Diag difference.
Are these caused by different processes? (This is trying to get at neutral evolution in cases of high Ne)
Anne's hypothesis: neutral morphological evolution (prediction: high Ne -> more diversity -> statistic differentiation but no diagnostic diff. in diverged species due to character overlap)
Anne's other hypothesis: stabilizing selection : high Ne -> efficient sel° -> + cases with « no Morpho differentiation ». 

# To test:

1. Let's look at HKK3. HKK3 is a proxy for Ne, where habitats are more or less connected. 

```{r cars}
surveymorpho<-filter(survey, Morpho_diff != "NA") #make matrix where only the cases where morpho was studied is here
Hkk_count<-table(surveymorpho$HKKv3)
morphodiff<-table(surveymorpho$Morpho_diff,surveymorpho$HKKv3) 
stat_diag<-morphodiff[3,]/morphodiff[1,]
stat_anydiff<-morphodiff[3,]/(morphodiff[3,]+morphodiff[1,])
stat_total<-morphodiff[3,]/Hkk_count
diag_total<-morphodiff[1,]/Hkk_count
nod_total<-morphodiff[2,]/Hkk_count
HKKtable<-rbind(stat_total,diag_total,nod_total)
morphodiff 
HKKtable
chisq.test(surveymorpho$Morpho_diff,surveymorpho$HKKv3)
```

Results:
- I think we need to ignore the first column (too few cases)
- Diagnostic characters show a decline as HKK increases - threshold between 25 and 100
- At large HKK, very few diagnostic traits


```{r}
HKKtable2<-pivot_longer(as.data.frame(HKKtable), cols=1:6, names_to = "HKK",
             values_to = "Percent", values_drop_na = FALSE)
HKKtable2$Diff_type<-c("Statistical","Statistical","Statistical","Statistical","Statistical","Statistical","Diagnostic","Diagnostic","Diagnostic","Diagnostic","Diagnostic","Diagnostic","No_Difference","No_Difference","No_Difference","No_Difference","No_Difference","No_Difference")
HKKtable2["HKK"][HKKtable2["HKK"] == "1"] <- "A1"
HKKtable2["HKK"][HKKtable2["HKK"] == "5"] <- "B2"
HKKtable2["HKK"][HKKtable2["HKK"] == "10"] <- "C10"
HKKtable2["HKK"][HKKtable2["HKK"] == "25"] <- "D25"
HKKtable2["HKK"][HKKtable2["HKK"] == "100"] <- "E100"
HKKtable2["HKK"][HKKtable2["HKK"] == "1000"] <- "F1000"

ggplot(as.data.frame(HKKtable2),aes(x=HKK,y=Percent,fill=Diff_type,))+geom_bar(stat="identity")

```

#2. Let's look at this by reproductive investment. (Toujours pas refait 6 oct)
(Idea from Romiguier et al. 2014, where genetic diversity is related to investment. They are looking at pi, but connect it verbally to Ne.)

First, I need to group the larval types into "large" and "small" reproductive investments.
(These data are still a mess, and I should clean them up before we do anything with the results, e.g. some are listed as 'viviparous'.) For now, if larvae are either planktotrophic or pelagic, I've classified as low investment. If lecithotrophic or dd, high.

```{r}
#Make dataframe of the species with larvae recorded
surveylarva<-filter(survey, larv.pktro != "NA") #this works because if NA, there's nothing for any of the larval columns

repro_invest<-c()

for (i in 1:length(surveylarva$NameBeforeCor)) {
  if (surveylarva[i,77] == TRUE)
    b<-"low"
  else if (surveylarva[i,80] == TRUE)
    b<- "low"
  else if (surveylarva[i,78] == TRUE)
    b<-"high"
  else if (surveylarva[i,79] == TRUE)
    b<-"high"
  else
    b<-NA
    
repro_invest<-c(repro_invest,b)
  
}

surveylarva<-cbind(surveylarva,repro_invest)

```

OK, let's do the tests now.

```{r}
table(surveylarva$repro_invest,surveylarva$Morpho_diff)
chisq.test(surveylarva$repro_invest,surveylarva$Morpho_diff)


```

For now (Oct 5), result is not significant. But I want to clean up the data and try again.


#3. Trying Morpho and Geo with sym/allo column removed

```{r}
surveymorpho<-filter(survey, Morpho_diff != "NA")
surveymorpho2<-filter(surveymorpho, Geo_dif3 != "Sympatric, Allopatric")
surveymorpho2<-droplevels(surveymorpho2)
#this removes cases where there are both options, n=12
#285 cases remaining where there is morpho data AND spp are either sympatric or allopatric

table(surveymorpho2$Morpho_diff,surveymorpho2$Geo_dif3)
chisq.test(surveymorpho2$Morpho_diff,surveymorpho2$Geo_dif3)


```

SO -- there is an association between morpho diff and geo diff.

```{r}
#and now a stacked bar
morphogeo_count<-table(surveymorpho2$Geo_dif3)
morphodiff<-table(surveymorpho2$Morpho_diff,surveymorpho2$Geo_dif3) 
stat_diag<-morphodiff[3,]/morphodiff[1,]
stat_anydiff<-morphodiff[3,]/(morphodiff[3,]+morphodiff[1,])
stat_total<-morphodiff[3,]/morphogeo_count
diag_total<-morphodiff[1,]/morphogeo_count
nod_total<-morphodiff[2,]/morphogeo_count
morphogeotable<-rbind(stat_total,diag_total,nod_total)
morphogeotable #this should be the percentages for each, allo and sympatric
#columns sum to 1

surveymorphogeotable<-pivot_longer(as.data.frame(morphogeotable), cols=1:2, names_to = "Geo_Diff",
             values_to = "Percent", values_drop_na = FALSE)
surveymorphogeotable$Diff_type<-c("Statistical","Statistical","Diagnostic","Diagnostic","No_Difference","No_Difference")

ggplot(as.data.frame(surveymorphogeotable),aes(x=Geo_Diff,y=Percent,fill=Diff_type,))+geom_bar(stat="identity")

```


For species in sympatry, it's more likely that they will be statistically differentiated. 

What about if I use NbCS per article = 1?
```{r}
surveymorpho3<-filter(surveymorpho2, Nb_CScomplex_inArticle == 1)
#filtering to remove large barcoding studies (etc)

table(surveymorpho3$Morpho_diff,surveymorpho3$Geo_dif3)
chisq.test(surveymorpho3$Morpho_diff,surveymorpho3$Geo_dif3)


```
Interestingly, when those studies are removed the effect goes away??

```{r}
morphogeo_count2<-table(surveymorpho3$Geo_dif3)
morphodiff2<-table(surveymorpho3$Morpho_diff,surveymorpho3$Geo_dif3) 
stat_diag2<-morphodiff2[3,]/morphodiff2[1,]
stat_anydiff2<-morphodiff2[3,]/(morphodiff2[3,]+morphodiff2[1,])
stat_total2<-morphodiff2[3,]/morphogeo_count2
diag_total2<-morphodiff2[1,]/morphogeo_count2
nod_total2<-morphodiff2[2,]/morphogeo_count2
morphogeotable2<-rbind(stat_total2,diag_total2,nod_total2)
morphogeotable2 #this should be the percentages for each, allo and sympatric
#columns sum to 1

surveymorphogeotable2<-pivot_longer(as.data.frame(morphogeotable2), cols=1:2, names_to = "Geo_Diff",
             values_to = "Percent", values_drop_na = FALSE)
surveymorphogeotable2$Diff_type<-c("Statistical","Statistical","Diagnostic","Diagnostic","No_Difference","No_Difference")

ggplot(as.data.frame(surveymorphogeotable2),aes(x=Geo_Diff,y=Percent,fill=Diff_type,))+geom_bar(stat="identity")

```
The patterns look broadly similar when the larger articles are removed, but it's no longer significant. Perhaps this has to do with sample size.

#4. Are there associations between Eco and Morpho diff if we only consider cases where RIOK?

```{r}
#First, start with cases where there is morphodiff
surveymorpho<-filter(survey, Morpho_diff != "NA") #make matrix where only the cases where morpho was studied is here
surveymorphoRIOK<-filter(surveymorpho, RIOK == TRUE) #make matrix where only the cases where RIOK = TRUE (good reproductive isolation)

table(surveymorphoRIOK$Morpho_diff,surveymorphoRIOK$Eco_diff)
chisq.test(surveymorphoRIOK$Morpho_diff,surveymorphoRIOK$Eco_diff)

#and a graph

morphoeco_count<-table(surveymorphoRIOK$Eco_diff)
morphodiff<-table(surveymorphoRIOK$Morpho_diff,surveymorphoRIOK$Eco_diff) 
stat_diag<-morphodiff[3,]/morphodiff[1,]
stat_anydiff<-morphodiff[3,]/(morphodiff[3,]+morphodiff[1,])
stat_total<-morphodiff[3,]/morphoeco_count
diag_total<-morphodiff[1,]/morphoeco_count
nod_total<-morphodiff[2,]/morphoeco_count
morphoecotable<-rbind(stat_total,diag_total,nod_total)
morphoecotable #this should be the percentages for each, eco and not
#columns sum to 1

surveymorphoecotable<-pivot_longer(as.data.frame(morphoecotable), cols=1:2, names_to = "Eco_Diff",
             values_to = "Percent", values_drop_na = FALSE)
surveymorphoecotable$Diff_type<-c("Statistical","Statistical","Diagnostic","Diagnostic","No_Difference","No_Difference")

ggplot(as.data.frame(surveymorphoecotable),aes(x=Eco_Diff,y=Percent,fill=Diff_type,))+geom_bar(stat="identity")

```


Result: No significant association between ecology and morphological differences when considering cases where RIOK = True. When there is ecological differentiation, a tendancy towards more diagnostic differences and fewer CSss, but not significant.

## Another series of questions - what biological factors correspond with CS
#6. Is skeleton type associated with morphological differentiation?

```{r}
#First, start with cases where there is morphodiff
surveymorpho<-filter(survey, Morpho_diff != "NA") #make matrix where only the cases where morpho was studied is here


table(surveymorpho$Morpho_diff,surveymorpho$hard_skeleton)
chisq.test(surveymorpho$Morpho_diff,surveymorpho$hard_skeleton)

#and a graph

morphoskel_count<-table(surveymorpho$hard_skeleton)
morphodiff<-table(surveymorpho$Morpho_diff,surveymorpho$hard_skeleton) 
stat_diag<-morphodiff[3,]/morphodiff[1,]
stat_anydiff<-morphodiff[3,]/(morphodiff[3,]+morphodiff[1,])
stat_total<-morphodiff[3,]/morphoskel_count
diag_total<-morphodiff[1,]/morphoskel_count
nod_total<-morphodiff[2,]/morphoskel_count
morphoskeltable<-rbind(stat_total,diag_total,nod_total)
morphoskeltable #this should be the percentages for each, eco and not
#columns sum to 1

surveymorphoskeltable<-pivot_longer(as.data.frame(morphoskeltable), cols=1:2, names_to = "Hard_Skeleton",
             values_to = "Percent", values_drop_na = FALSE)
surveymorphoskeltable$Diff_type<-c("Statistical","Statistical","Diagnostic","Diagnostic","No_Difference","No_Difference")

ggplot(as.data.frame(surveymorphoskeltable),aes(x=Hard_Skeleton,y=Percent,fill=Diff_type,))+geom_bar(stat="identity")

```

Result: This is significant. More No_diff if there's no hard skeleton, which is what we expect.

#7. Is image vision associated with morphological differentiation?

```{r}
#First, start with cases where there is morphodiff
surveymorpho<-filter(survey, Morpho_diff != "NA") #make matrix where only the cases where morpho was studied is here


table(surveymorpho$Morpho_diff,surveymorpho$image)
chisq.test(surveymorpho$Morpho_diff,surveymorpho$image)

```

Result: No, not significant.

#8. Are external genitals associated with morphological differentiation?

```{r}
#First, start with cases where there is morphodiff
surveymorpho<-filter(survey, Morpho_diff != "NA") #make matrix where only the cases where morpho was studied is here


table(surveymorpho$Morpho_diff,surveymorpho$genitals)
chisq.test(surveymorpho$Morpho_diff,surveymorpho$genitals)

#and a graph

morphogen_count<-table(surveymorpho$genitals)
morphodiff<-table(surveymorpho$Morpho_diff,surveymorpho$genitals) 
stat_diag<-morphodiff[3,]/morphodiff[1,]
stat_anydiff<-morphodiff[3,]/(morphodiff[3,]+morphodiff[1,])
stat_total<-morphodiff[3,]/morphogen_count
diag_total<-morphodiff[1,]/morphogen_count
nod_total<-morphodiff[2,]/morphogen_count
morphogentable<-rbind(stat_total,diag_total,nod_total)
morphogentable #this should be the percentages for each, eco and not
#columns sum to 1

surveymorphogentable<-pivot_longer(as.data.frame(morphogentable), cols=1:3, names_to = "External_genitals",
             values_to = "Percent", values_drop_na = FALSE)
surveymorphogentable$Diff_type<-c("Statistical","Statistical","Statistical","Diagnostic","Diagnostic","Diagnostic","No_Difference","No_Difference","No_Difference")

ggplot(as.data.frame(surveymorphogentable),aes(x=External_genitals,y=Percent,fill=Diff_type,))+geom_bar(stat="identity")

```

Result: Significant. Less statistical differentiation for "both" and less diagnostic for "no external genitals." 

Turns out that only gastropods are "both," so what happens if we remove them?

```{r}
#First, start with cases where there is morphodiff
surveymorpho<-filter(survey, Morpho_diff != "NA") #make matrix where only the cases where morpho was studied is here
surveymorphonob<-filter(surveymorpho, genitals != "B")
surveymorphonob<-droplevels(surveymorphonob)


table(surveymorphonob$Morpho_diff,surveymorphonob$genitals)
chisq.test(surveymorphonob$Morpho_diff,surveymorphonob$genitals)

#and a graph

morphogen_count<-table(surveymorphonob$genitals)
morphodiff<-table(surveymorphonob$Morpho_diff,surveymorphonob$genitals) 
stat_diag<-morphodiff[3,]/morphodiff[1,]
stat_anydiff<-morphodiff[3,]/(morphodiff[3,]+morphodiff[1,])
stat_total<-morphodiff[3,]/morphogen_count
diag_total<-morphodiff[1,]/morphogen_count
nod_total<-morphodiff[2,]/morphogen_count
morphogentable<-rbind(stat_total,diag_total,nod_total)
morphogentable #this should be the percentages for each, eco and not
#columns sum to 1

surveymorphonobgentable<-pivot_longer(as.data.frame(morphogentable), cols=1:2, names_to = "External_genitals",
             values_to = "Percent", values_drop_na = FALSE)
surveymorphonobgentable$Diff_type<-c("Statistical","Statistical","Diagnostic","Diagnostic","No_Difference","No_Difference")

ggplot(as.data.frame(surveymorphonobgentable),aes(x=External_genitals,y=Percent,fill=Diff_type,))+geom_bar(stat="identity")

```

Result: Still significant: without external genitals, more likely to have CSss.